A Prediction of the 2024 U.S. Election Outcome 

### 2024-11-04 paper/paper.qmd
Add files via upload
Gramma spelling.

## Predictor Variable: State
```{r}
#| label: state_dist
#| fig-cap: Distribution of Observations by State
#| echo: false
analysis_data$state[is.na(analysis_data$state)] <- "National"
state_counts <- analysis_data %>%
  count(state) %>%
  mutate(percentage = n / sum(n) * 100)
ggplot(state_counts, aes(x = state, y = percentage)) +
  geom_bar(stat = "identity", color = "black", fill = "lightblue") +
  labs(title = "Distribution of Observations by State (Percentage)", x = "State", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The state where the polling took place. This is important for state-level fixed effects, some regions show more supoprt to one party over another. Regarding the data itself, ee notice that over half of all observations are NA, in our analysis we take these to be national level polls. Since this is not explicit in the documentation [@538], this is an assumption which is a caveat in our outcome analysis. We notice in the figure (@state_dist) that not every state is represented, which limits us to the states we can make state-level predictions for.

@@

##Outcome Variable: pct

@@

- **Comparative Insights**: By comparing the two we find that the Democrats are in the lead but not by a significant margin, and that the Republicans are showing some recent momentum. However, the credibility intervals are so large, they overlap. As such, there is no significant conclusion. Within the credibility intervals are scenarios where the Republicans lose with 42% and the Democrats win with 52%, as well as scenarios where the Republicans win with 51% and the Democrats lose with 42%.

## Posterior predictive check

@@

Our results section allows us to make some inferences. First, we see that the Democrats are ahead, and we predict a win for Kamala Harris. However, the time series of the spline shows us that their support has been flat over time, whereas the Republican party has been showing some recent momentum which could contribute to a surprise. This is in-line with some of the literatue, with @cohn_2024 showing how Trump brings his strength later in the election cycle as voters are generally shy to show their support for him.

@@

We can conclude, that the democrats are strongly favoured to win, but the Republicans do also have a stronger probability of an upward surprise, whereas the democrats have a stronger probabiltiy of a downward surprise.

@@

### Budgect Allocation

>>>>

## Predictor Variable: State
```{r}
#| label: state_dist
#| fig-cap: Distribution of Observations by State
#| echo: false
analysis_data$state[is.na(analysis_data$state)] <- "National"
state_counts <- analysis_data %>%
  count(state) %>%
  mutate(percentage = n / sum(n) * 100)
ggplot(state_counts, aes(x = state, y = percentage)) +
  geom_bar(stat = "identity", color = "black", fill = "lightblue") +
  labs(title = "Distribution of Observations by State (Percentage)", x = "State", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The state where the polling took place. This is important for state-level fixed effects, some regions show more support to one party over another. Regarding the data itself, we notice that over half of all observations are NA, in our analysis we take these to be national level polls. Since this is not explicit in the documentation [@538], this is an assumption which is a caveat in our outcome analysis. We notice in the figure (@state_dist) that not every state is represented, which limits us to the states we can make state-level predictions for.

@@

## Outcome Variable: pct

@@

- **Comparative Insights**: By comparing the two we find that the Democrats are in the lead but not by a significant margin, and that the Republicans are showing some recent momentum. However, the credibility intervals are so large, they overlap. As such, there is no statistically significant conclusion. Within the credibility intervals are scenarios where the Republicans lose with 42% and the Democrats win with 52%, as well as scenarios where the Republicans win with 51% and the Democrats lose with 42%.

## Posterior Predictive check

@@

Our results section allows us to make some inferences. First, we see that the Democrats are ahead, and we predict a win for Kamala Harris. However, the time series of the spline shows us that their support has been flat over time, whereas the Republican party has been showing some recent momentum which could contribute to a surprise. This is in-line with some of the literature, with @cohn_2024 showing how Trump brings his strength later in the election cycle as voters are generally shy to show their support for him.

@@

We can conclude, that the democrats are strongly favored to win, but the Republicans do also have a stronger probability of an upward surprise, whereas the democrats have a stronger probability of a downward surprise.

@@

### Budget Allocation
